---
title: "Research Practices"
subtitle: "Organization, Tools, and Reproducibility"
author: "Chuck Lanfear"
date: 'Oct 25, 2018'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
csss508css <- list(
     ".hljs-tomorrow-night-bright .hljs" = list(
        "background" = "#10102c",
        "border-radius"="5px"),
     ".remark-inline-code" = list(
        "background" = "#E0E0E0",
        "color" = "#10102c",
        "border-radius" = "3px",
        "padding" = "2px"),
     ".inverse .remark-inline-code" = list(
        "background" = "#10102c",
        "color" = "#ececf8",
        "border-radius" = "3px",
        "padding" = "2px"),
     ".small" = list("font-size" = "75%"),
     ".smallish" = list("font-size" = "85%"),
     ".remark-code-line-highlighted" = list(
        "background-color" = "rgba(255,0,255,0.2)"),
     "sup" = list("font-size" = "14px"),
     "h1, h2, h3" = list(
        "margin-top"=".25em", 
        "margin-bottom"=".25em"),
     ".title-slide h1" = list(
       "margin-top" = "-180px"
     ),
     ".pull-left60" = list(
        "float" = "left",
        "width" = "58%" ),
     ".pull-right40" = list(
        "float" = "right",
        "width" = "38%" ),
     ".pull-right30" = list(
        "float" = "right",
        "width" = "30%" ),
     "a" = list("text-decoration" = "underline"),
     ".inverse a" = list("color" = "#cbd3a3"),
     "body" = list("line-height" = "1.4"),
     ".inverse" = list("background-image" = "url(https://clanfear.github.io/BE_102518/img/UWSocBG1_white.svg)"),
     ".centernote" = list(
        "text-align" = "center",
        "position" = "fixed",
        "left" = "0",
        "right" = "0",
        "margin" = "0 auto",
        "padding-left" = "100px",
        "padding-right" = "100px",
        "bottom" = "3em"),
     ".image-50 img" = list(
        "width" = "50%",
        "display"="block",
        "margin-left"="auto",
        "margin-right"="auto"),
     ".image-75 img" = list(
        "width" = "75%",
        "display"="block",
        "margin-left"="auto",
        "margin-right"="auto"),
     ".image-full img" = list(
        "height" = "480px",
        "display"="block",
        "margin-left"="auto",
        "margin-right"="auto"),
     ".image-center img" = list(
        "display"="block",
        "margin-left"="auto",
        "margin-right"="auto"),
     ".center table" = list("display" = "inline-block"),
     ".title-slide" = list("background-image" = "url(https://clanfear.github.io/BE_102518/img/UWSocBG1_white.svg), url(https://clanfear.github.io/SocSem_102618/img/title_slide_img.png)",
                           "background-position" = "center center, center bottom",
                           "background-size" = "contain, contain"),
     ".remark-slide thead, .remark-slide tfoot, .remark-slide tr:nth-child(even)" = list("background" = "#FFFFFF"),
     ".remark-slide table th" = list("border-bottom" = "1px solid #000",
                                     "border-top" = "4px double #000",
                                     "text-align" = "center",
                                     "font-weight" = "normal",
                                     "font-style" ="italic"),
     ".remark-slide table" = list("border-top" = "none",
                                  "border-bottom" = "1px solid #000",
                                  "border-collapse" = "collapse"),
     ".remark-slide table tr" = list(
       "line-height" = "20px"
     ),
     ".remark-slide table td" = list(
       "padding" = "2px",
       "text-align" = "center"
     ),
     "strong" = list("color" = "#905145"),
     ".table-title" = list("font-size" = "115%")
    )
mono_accent(base_color = "#342c5c",
            code_font_google = google_font("Fira Mono"),
            header_font_google = google_font("Quattrocento"),
            extra_css = csss508css,
            background_image = "https://clanfear.github.io/BE_102518/img/UWSocBG1.svg"
)
```

# Overview

### 1. Organization
### 2. Tools and Skills
### 3. Reproducible Research
### 4. Version Control

---
class: inverse
# Organization

---
# A System

Developing an organization system for research is something you either do accidentally—and badly—or purposefully with some upfront labor.

--

Every paper, project, or presentation should be organized the same way.

--

I suggest something like the following (an R example):

.pull-left[
```
project/
   readme.md
   data/
     derived/
       processed_data.RData
     raw/
       core_data.csv
   docs/
     paper.Rmd
   scripts/
     functions.R
     models.R
```
]
.pull-right[
1. There is a clear hierarchy
   * Written content is in `docs`
   * Code is in `scripts`
   * Data is in `data`
2. Naming is uniform
   * All lower case
   * Words separated by underscores
3. Names are self-descriptive

]

---
class: inverse
# Tools

---
# On Formats

Avoid "closed" or commercial software and file formats except where absolutely necessary.

--

Use open source software and file formats.

--

* It is always better for *science*:

   + People should be able to explore your research without buying commercial software.
   + You do not want your research to be inaccessible when software is updated.

--

* It is often just *better*.

   + It is usually updated more quickly
   + It tends to be more secure
   + It is rarely abandoned

--

**The ideal:** Use software that reads and writes *raw text*.

---
# Programming

Learn to program.

--

Coding is a universal, transferrable, and marketable skill.

--

People are bad at repetitive tasks, but computers are very good.

--

Programming is telling a computer how to do repetitive tasks for you.

--

If what you do can be automated easily, it is not a useful skill—*but knowing how to automate it is*.

--

Most research labor is easily automated, *but people do it manually anyway.*

**Stop.**

---
# R or Python

R and Python dominate the general sphere of scientific computing.

--

.pull-left[
### R

* Older, more idiosyncratic
* Dominant in academia
* Best tools for inference
* Superior document authoring
]

--

.pull-right[
### Python

* Newer with simpler syntax
* Dominant in industry
* Best tools for prediction
* Easiest to deploy at scale
]

--

*There is nothing one can do that the other can't<sup>1</sup> and they are comparable in performance<sup>2</sup>, so learn one well and branch out later*

.footnote[
[1] Though it may be *harder*!

[2] If you write code correctly.
]

---
# Getting Started

Setting up and using a programming environment for the first time can be daunting. I recommend using common platforms for either.

--

.pull-left[
### R

1. [Get R](https://cran.r-project.org/)
2. [Get RStudio](https://www.rstudio.com/products/rstudio/download/#download)
3. Find a tutorial
   * I like [DataCamp](https://www.datacamp.com/)
   * [swirl](https://swirlstats.com/) is great too
4. Take a class
   * [I like mine, I teach it](https://clanfear.github.io/CSSS508/)

]

--

.pull-right[
### Python

1. [Get Python 3](https://www.python.org/downloads/)
2. Get an Integrated Development Environment (IDE) <sup>1</sup>
   * [I use Atom](https://atom.io/)<sup>2</sup>
   * [PyCharm is good too](https://www.jetbrains.com/pycharm/)
3. Find a tutorial
   * I like [DataCamp](https://www.datacamp.com/)
4. Take a class
   * CSE583 is great
]

.footnote[
[1] [Jupyter Notebooks are kind of an IDE](http://jupyter.org/)

[2] Atom is an excellent *general text editor*. I use it for *everything*.
]
---
# Text

Writing and formatting documents are two completely separate jobs.
   * Write first
   * Format later
   * [Markdown](https://en.wikipedia.org/wiki/Markdown) was made for this

--

Word processors—like Microsoft Word—try to do both at the same time, usually badly.

They waste time by leading you to format instead of writing.

--

Find a good modular text editor and learn to use it:
   * [Atom](https://atom.io/)
   * [Sublime](https://www.sublimetext.com/) (Commercial)
   * Vim<sup>1</sup>
   * Emacs<sup>1</sup>

.footnote[[1] These are powerful but complex. RStudio has a Vim mode too.]

---
class: inverse

# Reproducible Research

---
## Reproducibility Definitions

Reproducibility comes in three forms (Stodden 2014):

--

1. **Empirical:** Repeatability in data collection.

--

2. **Statistical:** Verification with alternate methods of inference.

--

3. **Computational:** Reproducibility in cleaning, organizing, and presenting data and results.

--

R and Python are particularly well suited to enabling **computational reproducibility**.

--

They will not fix flawed experimental methods or observations, nor offer a remedy for improper application of statistical methods.

Those are the difficult, non-repetitive things you want skills in.

---

## Computational Reproducibility

Elements of computational reproducibility:
* Shared data
* Shared code
* Documentation
* **Version Control**

---

## Levels of Reproducibility

For academic papers, degrees of reproducibility vary:

0. "Read the article"
--

1. Shared data with documentation
--

2. Shared data and all code
--

3. **Interactive document**
--

4. **Research compendium**
--

5. Docker File: Self-contained ecosystem

---
## Interactive Documents

Interactive documents combine code and text together
   * Loading and process data
   * Run models
   * Generate tables and plots in-line with text
   * In-text values automatically filled in

--

Interactive documents allow a reader to examine your computational methods within the document itself.

--

By re-running the code, they reproduce your results on demand.

--

Platforms:
 * **R:** RMarkdown ([an example of mine](https://clanfear.github.io/birthtiming/inst/paper/paper.html))
 * **Python:** Jupyter Notebooks

---
## Research Compendia

A research compendium is an R package used not to share statistical or computational *methods* but to organize and share reproducible *projects*.

--

Research compendia feature:

* Files organized in a recognizable structure (R packages)
* Clear separation of data, method, and output. *Data are read only*.
* A well-documented or even *preserved* computational environment

`rrtools` by UW's Ben Markwick provides a simplified workflow to accomplish this in R.

---
## Bookdown

If using your compendium to generate an article, thesis, or dissertation, `rrtools` makes formatting simple by integrating `bookdown`.

`bookdown` provides an accessible alternative to manually writing $\LaTeX$ for typesetting and reference management.

You can integrate citations and automate reference page generation using bibtex files (such as produced by Zotero).

`bookdown` supports `.html` output for ease and speed and also renders `.pdf` files through $\LaTeX$ for publication-ready documents.

For University of Washington theses and dissertations, consider Ben Marwick's `huskydown` package which uses Markdown but renders via a UW approved $\LaTeX$ template.

---
# Version Control

Version control originates in collaborative software development.

**The Idea:** All changes ever made to a piece of software are documented, saved automatically, and revertible.

--

Version control allows all decisions ever made in a research project to be documented automatically.

--

Version control can:

1. Protect your work from destructive changes
2. Simplify collaboration by merging changes
3. Document design decisions
4. Make your research process transparent

---
# Git and GitHub

[Git](https://en.wikipedia.org/wiki/Git) is the dominant platform for version control, and [GitHub](https://github.com/) is a (now Microsoft owned) platform for hosting **repositories**.

--

Repositories are folders on your computer where all changes are tracked by Git.

--

Once satisfied with changes, you "commit" them then "push" them to a remote repository that stores your project.

--

Others can copy your project ("pull"), and if you permit, make suggestions for changes.

--

Constantly committing and pulling changes automatically generates a running "history" that documents the evolution of a project.

---
# Summary

Good research hygiene is good for science

* Spend labor up-front to reduce headaches later on

* Learn to program

   * Replace boring things with automation
   * Spend your time on hard problems instead

* Do good science

   * Reproducible
   * Accessible
   * Transparent
